---
title: "Understanding HIV Viral Load After Treatment"
output: 
  knitrBootstrap::bootstrap_document:
    title: "Understanding HIV Viral Load After Treatment"
    theme: united
    highlight: zenburn
    theme.chooser: TRUE
    highlight.chooser: TRUE
---

# Data Loading and Transformation

## Grabbing the Raw Data

```{r setup, include=TRUE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(magrittr)
library(ggplot2)
library(dplyr)
library(knitr)

hiv_data <- read_csv(
  "https://raw.githubusercontent.com/ComputationalModeling/IPML-Data/master/01HIVseries/HIVseries.csv",
  col_names = c("time_in_days", "viral_load")
)
```

## Log-transforming the Viral Load Data 

```{r}
hiv_data %<>%
  mutate(log_viral_load = log(viral_load)) %>%
  mutate(index = rownames(hiv_data))

hiv_data %>% knitr::kable()
```

# Defining useful helper functions

```{r}
time <- seq(from = 0, to = 10, length.out = 101)

get_model <- function(A, alpha, B, beta, time) {
  model <- (A * exp(-alpha * time)) + (B * exp(-beta * time))
  return (model)
}

get_slope_and_intercept <- function(
  x_values,
  y_values,
  index1,
  index2)
{
  rise <-  y_values[index2] - y_values[index1]
  run <-  x_values[index2] - x_values[1]
  slope <-  rise / run
  intercept <-  y_values[index1] - (slope * x_values[index1])

  output <- data.frame(
    rise,
    run,
    slope,
    intercept
  )
  return (output)
}
```

# Data Analysis

## Analytically determining the slopes and intercepts of two log-linear models for virion clearance

```{r}
early_stage_clearance <- get_slope_and_intercept(
  x_values = hiv_data$time_in_days,
  y_values = hiv_data$log_viral_load,
  index1 = 1,
  index2 = 10
) %>% 
  data.frame(clearance_stage="early stage")

late_stage_clearance <- get_slope_and_intercept(
  x_values = hiv_data$time_in_days,
  y_values = hiv_data$log_viral_load,
  index1 = 11,
  index2 = 16
) %>% 
  data.frame(clearance_stage="later stage")

clearance_rates <- merge(
  early_stage_clearance, 
  late_stage_clearance,
  all = TRUE
)

clearance_rates %>% kable()
```

## Plotting the differing linear models for virion clearance

We can try plotting the two distinct linear models

```{r}
raw_data_plot <- ggplot()
raw_data_plot <- raw_data_plot + geom_text(
  aes(
    x = time_in_days,
    y = log_viral_load,
    label = index
  ),
  data = hiv_data
)

full_line_plot <- geom_abline(
  aes(
    slope = slope,
    intercept = intercept,
    group = clearance_stage,
    color = clearance_stage
  ),
  data = clearance_rates
)

print(
  raw_data_plot +
    full_line_plot + 
    scale_color_brewer(type = "qual")
)

```

## Plotting the differing linear models as simply line segments


```{r}
clearance_1 <- with(
  hiv_data,
  data.frame(
    x = time_in_days[1],
    xend = time_in_days[10],
    y = log_viral_load[1],
    yend = log_viral_load[10],
    clearance_stage = "early"
  )
)

clearance_2 <- with(
  hiv_data,
  data.frame(
    x = time_in_days[11],
    xend = time_in_days[16],
    y = log_viral_load[11],
    yend = log_viral_load[16],
    clearance_stage = "late"
  )
)

clearance_rates <- merge(
  clearance_1,
  clearance_2,
  all = TRUE
)

line_segment_plot <- raw_data_plot
line_segment_plot <- line_segment_plot + geom_segment(
  aes(
    x = x,
    xend = xend,
    y = y,
    yend = yend,
    group = clearance_stage,
    color = clearance_stage
  ),
  data = clearance_rates
)

line_segment_plot <- line_segment_plot + 
  scale_color_brewer(type = "qual") +
  theme_bw()

print(line_segment_plot)
```

)

